---
title: "Assessment_Cleo_Schwarz"
author: Cleo Schwarz
date: "`r date()`"
format: 
  html:
    toc: true
    code-fold: show
    code-tools: true
    df-print: paged
    embed-resources: true 
---

```{r}
#| include: false 
# Load libraries
library(here) 
library(tidyverse) 
library(lubridate) 
library(gtsummary)
library(rstatix)
```

## Introduction

Smokers as well as a high BMI are both well known risk factors for any disease. In this analysis the focus lays on smokers.

The aim of this study was to investigate whether smoking has an impact on the individual insurance charges. 

## Data

### Read data and transform data

The data for this analysis is open source and was obtained from [Kaggel](https://www.kaggle.com/datasets/mirichoi0218/insurance?resource=download). From all data, only the values of interest were selected (e.g. age, bmi, sex, smoker, charges). 

```{r}
#| output: false
dat <- read_csv(here("data/raw/insurance_with_date.csv"))
str(dat)

insurance <- dat |> 
  select(X, age, sex, bmi, smoker, charges) |> 
  mutate(sex = factor(sex, levels=c("male", "female"), labels=c("m","f")), 
         smoker = factor(smoker))
```

### Descriptive statistics

This table shows the mean and median, SD (standard deviation), smallest and largest values of charges, age and bmi sorted by smorkers.

```{r}
summarysmoke <- insurance |>
  tbl_summary(
    by = smoker, 
    type = all_continuous() ~ "continuous2", 
    include =c("charges","bmi", "age"),
    statistic = all_continuous() ~ c("{mean} ({sd})",
                                  "{median} ({p25}, {p75})",
                                  "{min}, {max}")
  ) 
```

### Test for normality

For visual check if the charges are in normal distribution a QQ-plot and a histogram was created. Further, the shapiro test was performed.

####QQ-Plot

```{r}
insurance %>%
  ggplot(aes(sample = charges, fill = smoker, colour = smoker)) + 
  geom_qq_line(distribution = stats::qnorm) +
  geom_qq(distribution = stats::qnorm) + 
  xlab("theoretical") +
  ylab("sample") +
  theme_classic() +
  facet_wrap("sex")
```

####Histogram

```{r}
ggplot(insurance, aes(x=charges)) +
  geom_histogram(bins=40) +
  theme_classic()
```

This analysis was also conducted by taking the logarithmic values of "charges"

####QQ-Plot - log

```{r}
insurance$charges_log <- log(insurance$charges)

insurance %>%
  ggplot(aes(sample = charges_log, fill = smoker, colour = smoker)) + 
  geom_qq_line(distribution = stats::qnorm) +
  geom_qq(distribution = stats::qnorm) + 
  xlab("theoretical") +
  ylab("sample") +
  theme_classic() +
  facet_wrap("smoker")

insurance |> 
  group_by(smoker) |> 
  shapiro_test(charges_log)
```

####Histogram - log

```{r}
ggplot(insurance, aes(x=charges_log)) +
  geom_histogram(bins=40) +
  theme_classic()
```

By doing the logarithm a conversion in normal distribution was possible.

## t-test

For comparison of the means of charges of smokers vs. non smokers a Wilcoxon Test was performed. The analysis was also conducted with the logarithmic values of charges by doing a paired ttest with them.

```{r}

insurance |> 
  t_test(charges_log ~ smoker) |> 
  rename(non_smoker = group1, smoker = group2)
         
  
insurance |> 
  wilcox_test(charges ~ smoker) 

```

# plots

```{r}
ggplot(insurance , aes(x = smoker, y = charges, fill = smoker )) + 
  geom_boxplot() +
  ggtitle(label = "Smokers vs. Charges") +
  xlab(label = "smoker") +
  ylab(label = "charges")+
  theme_classic() + theme(legend.position="bottom")
```

## Results

## Conclusion

There is a statistical significance between higher charges and smoking.
