---
title: "Assessment_Cleo_Schwarz"
author: Cleo Schwarz
date: "`r date()`"
format: 
  html:
    toc: true
    code-fold: show
    code-tools: true
    df-print: paged
    embed-resources: true 
---

```{r}
#| include: false 
# Load libraries
library(here) 
library(tidyverse) 
library(lubridate) 
library(gtsummary)
library(rstatix)
```

## Introduction

Smokers as well as a high bmi are both well known risk factors for any disease. 
For this analysis we focused on one of these risk factors, smoking. 

The aim of this analysis was to investigate whether smoking has an impact on insurance charges or in other words if smokers pay more insurance charges.

## Data

### Read data and transform data

The data for this analysis was obtained from kaggle. 
The dataset was selected for only the important variables. 

```{r}
#| output: false
dat <- read_csv(here("data/raw/insurance_with_date.csv"))
str(dat)

insurance <- dat |> 
  select(X, age, sex, bmi, smoker, charges) |> 
  mutate(sex = factor(sex, levels=c("male", "female"), labels=c("m","f")), 
         smoker = factor(smoker))
```

### Descriptive statistics

This table shows the mean and median, SD (standard deviation), smallest and largest values of charges, age and bmi sorted by smorkers.

```{r}
summarysmoke <- insurance |>
  tbl_summary(
    by = smoker, 
    type = all_continuous() ~ "continuous2", 
    include =c("charges","bmi", "age"),
    statistic = all_continuous() ~ c("{mean} ({sd})",
                                  "{median} ({p25}, {p75})",
                                  "{min}, {max}")
  ) 
```

### Test for normality

For visual check if the charges are in normal distribution a QQ-plot and a histogram was created. Further, the shapiro test was performed. 

####QQ-Plot

```{r}
insurance %>%
  ggplot(aes(sample = charges, fill = smoker, colour = smoker)) + 
  geom_qq_line(distribution = stats::qnorm) +
  geom_qq(distribution = stats::qnorm) + 
  xlab("theoretical") +
  ylab("sample") +
  theme_classic() +
  facet_wrap("sex")
```

####Histogram

```{r}
ggplot(insurance, aes(x=charges)) +
  geom_histogram(bins=40) +
  theme_classic()
```

This analysis was also conducted by taking the logarithmic values of "charges"

####QQ-Plot - log

```{r}
insurance$charges_log <- log(insurance$charges)

insurance %>%
  ggplot(aes(sample = charges_log, fill = smoker, colour = smoker)) + 
  geom_qq_line(distribution = stats::qnorm) +
  geom_qq(distribution = stats::qnorm) + 
  xlab("theoretical") +
  ylab("sample") +
  theme_classic() +
  facet_wrap("smoker")

insurance |> 
  group_by(smoker) |> 
  shapiro_test(charges_log)
```

####Histogram - log

```{r}
ggplot(insurance, aes(x=charges_log)) +
  geom_histogram(bins=40) +
  theme_classic()
```

By doing the logarithm a conversion in normal distribution was possible.

## t-test

For comparison of the means of charges of smokers vs. non smokers a Wilcoxon Test was performed. The analysis was also conducted with the logarithmic values of charges by doing a paired ttest with them.

```{r}

insurance |> 
  t_test(charges_log ~ smoker) |> 
  rename(non_smoker = group1, smoker = group2)
         
  
insurance |> 
  wilcox_test(charges ~ smoker) 

```

## plots

### boxplot 1 smokers vs charges

ggplot(insurance , aes(x = smoker, y = charges, fill = smoker )) + geom_boxplot() + ggtitle(label = "Smokers vs. Charges") + xlab(label = "smoker") + ylab(label = "charges")+ theme_classic() + theme(legend.position="bottom")

# plots 
```{r}
ggplot(insurance , aes(x = smoker, y = charges, fill = smoker )) + 
  geom_boxplot() +
  ggtitle(label = "Smokers vs. Charges") +
  xlab(label = "smoker") +
  ylab(label = "charges")+
  theme_classic() + theme(legend.position="bottom")
```


## Results

## Conclusion

There is a statistical significance between higher charges and smoking.
