---
title: "Ebola in Guinea, Liberia and Sierra Leone"
author: "Cleo Schwarz"
date: "r date()"
format: 
  html:
    toc: true
    code-fold: show
    code-tools: true
    df-print: paged
    embed-resources: true 
---

```{r}
#| include: false 
# Load libraries
library(here) 
library(tidyverse) 
library(lubridate) 
library(unibeCols)
```

## Introduction

The aim of this exercise is to visualize Ebola outbreaks in Guinea, Liberia and Sierra Leone.

## Data

Load data from data.world 

```{r}
# read data

cases <- read_csv("C:/Users/cleos/OneDrive/Universität Bern/Kurse/R_project_template/data/raw/ebola.csv") 


```

```{r}
```

```{r}
```

```{r}
# Process data
selection <- c("Sierra Leone","Guinea","Liberia")
window_start <- ymd("2014-08-28")
window_end <- ymd("2014-04-01")
cases_selection <- cases |> 
  filter(Country %in% selection & Date >= window_start & Date <= window_end) |>
  select(Date, Cum_conf_cases)
```

## Results 
@fig-cases shows the number of Ebola cases in Guinea, Sierra Leone and Liberia from `r window_start`to `r window_end`

```{r}
#| label: fig-cases
#| fig-cap: "Ebola Cases in Guinea, Sierra Leone and Liberia in 2014" 
#| fig-width: 8
#| fig-height: 4
ggplot(cases, aes(x = Date, y = Cum_conf_cases)) +
  geom_bar(stat = "identity", fill = unibeSaphireS()[1]) +
  labs(x = "", y ="Ebola cases", ) +
  theme_minimal() 
```

# Conclusions

The number of Ebola cases in Guniea, Sierra Leone and Liberia rapidly increased considering cases before March 31th, 2015. 
